Two Phase Debiased Calibration Simulation

### 2024-04-27 src/explore/20240411-tpdcsim.qmd
Updated Topic 1 simulation results. We are using a SRS-Poisson setup.

@@

@@

  # Variance Estimation
  ex_var_mat <- matrix(nrow = 2, ncol = 2)
  ex_var_mat[1, 1] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1^2)
  ex_var_mat[1, 2] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1 * p1_df$X2)
  ex_var_mat[2, 1] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1 * p1_df$X2)
  ex_var_mat[2, 2] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X2^2)

  Pi_dc <- diag(1 - 1 / d_vec) * (dc_w / pop_obs)^2
  z_dc <- model.matrix(~1 + p2_df$X1 + p2_df$X2 + gdi)
  gam_dc <- 
    solve(t(z_dc) %*% diag(gpinv) %*% z_dc, t(z_dc) %*% diag(gpinv) %*% p2_df$Y)
  vhat <- 
  (p2_df$Y - t(gam_dc) %*% t(z_dc)) %*% Pi_dc %*% t(p2_df$Y - t(gam_dc) %*% t(z_dc)) +
  (gam_dc[2:3] %*% ex_var_mat %*% gam_dc[2:3] / pop_obs^2)

  # FIXME: Change variance estimation to linearization estimate

@@

  pistar <- pi_star(p1_df, p2_df, nrow(pop_df), pop_df = pop_df)
  tpreg <- tp_reg(p1_df, p2_df, nrow(pop_df))
  # FIXME: GEcalib is not working. It produces results with very high
  # values of the function.
  dcpop <- 
    dc_poptots(
      p1_df,
      p2_df,
      nrow(pop_df),
      pop_tot = c(nrow(pop_df), sum(pop_df$X1), sum(pop_df$X2), sum(-pop_pi1))
    )
  dcest <-
    dc_esttots(
      p1_df,
      p2_df,
      nrow(pop_df),
      pop_gdi = sum(-pop_pi1)
    )

@@

>>>>

  # TODO: I am not sure if d_vec is correct
  d_vec <- 1 / (p2_df$pi2 * p2_df$pi1)
  if (entropy == "EL") {
    gdi <- -1 / d_vec
    gpinv <- d_vec^2

@@

  if (any(is.na(dc_w))) {
    return(list(theta = NA, Var = NA))
  }

@@

  # Variance Estimation
  # ex_var_mat <- matrix(nrow = 2, ncol = 2)
  # ex_var_mat[1, 1] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1^2)
  # ex_var_mat[1, 2] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1 * p1_df$X2)
  # ex_var_mat[2, 1] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X1 * p1_df$X2)
  # ex_var_mat[2, 2] <- sum((1 / p1_df$pi1^2 - 1 / p1_df$pi1) * p1_df$X2^2)

  # Pi_dc <- diag(1 - 1 / d_vec) * (dc_w / pop_obs)^2
  z_dc <- model.matrix(~1 + p2_df$X1 + p2_df$X2 + gdi)
  # gam_dc <- 
  #   solve(t(z_dc) %*% diag(gpinv) %*% z_dc, t(z_dc) %*% diag(gpinv) %*% p2_df$Y)
  # vhat <- 
  # (p2_df$Y - t(gam_dc) %*% t(z_dc)) %*% Pi_dc %*% t(p2_df$Y - t(gam_dc) %*% t(z_dc)) +
  # (gam_dc[2:3] %*% ex_var_mat %*% gam_dc[2:3] / pop_obs^2)

  phi_est <- 
    base::solve(t(z_dc) %*% diag(gpinv) %*% z_dc, t(z_dc) %*% (p2_df$Y * gpinv))

  p12_df <- 
    left_join(p1_df, p2_df, by = join_by(X1, X2, X3, X4, Y, pi1, pi2, del1)) %>%
    mutate(del2 = ifelse(is.na(del2), 0, 1))

  pred12 <- 
  model.matrix(~1 + p12_df$X1 + p12_df$X2 + I(-1/(p12_df$pi2 + p12_df$pi1))) %*%
    phi_est

  model.matrix(~1 + p12_df$X1 + p12_df$X2 + I(-1/(p12_df$pi2 + p12_df$pi1))) %>%
    head
  pred2 <- 
  model.matrix(~1 + p2_df$X1 + p2_df$X2 + I(-1/(p2_df$pi2 + p2_df$pi1))) %*%
    phi_est
  eta <- pred12 + p12_df$del2 / p12_df$pi2 * (p12_df$Y - pred12)
  v1 <- sum((1 - p12_df$pi1) / p12_df$pi1^2 * eta^2)
  v2 <- sum(1 / (p2_df$pi1 * p2_df$pi2) * (1 / p2_df$pi2 - 1) * (p2_df$Y - pred2)^2)
  vhat <- (v1 + v2) / pop_obs^2

@@

  pistar <- pi_star(p1_df, p2_df, nrow(pop_df), pop_df = pop_df)
  tpreg <- tp_reg(p1_df, p2_df, nrow(pop_df))
  dcpop <- 
    dc_poptots(
      p1_df,
      p2_df,
      nrow(pop_df),
      pop_tot = c(nrow(pop_df), sum(pop_df$X1), sum(pop_df$X2), sum(-pop_pi1))
    )
  dcest <-
    dc_esttots(
      p1_df,
      p2_df,
      nrow(pop_df),
      pop_gdi = sum(-pop_pi1)
    )

@@

mc_res %>%
  filter(Est == "dcpop") %>%
  arrange(desc(Theta))
